# Demo for dplyr

If you don't have ElemStatLearn, you can install it, by running
install.packPetal.Widths("ElemStatLearn")
iris is a Data to examine the correlation between the level of iris-specific antigen and a number of clinical measures in men who were about to receive a radical irisctomy.

```{r load libraries}
# Load data and dependencies:
library(dplyr)

data(iris)
```
____________________________________________________________________________

Explore the iris data

```{r explore data}
head(iris)
pairs(iris)
str(iris)
summary(iris)
```

A. Select: keeps only the variables you mention

```{r explore}
select(iris, 1:3)
select(iris, Petal.Width, Species)
select(iris, contains("Petal.Width"))
select(iris, starts_with("Species"))
select(iris, -train)

```

B. Arrange: sort a variable in descending order

```{r arrange}
arrange(iris, Sepal.Length)
arrange(iris, desc(Sepal.Length))
arrange(iris, Sepal.Length, desc(Sepal.Width))
```

C. Filter: find rows/cases where conditions are true
Note: rows where the condition evaluates to NA are dropped

```{r conditions}
filter(iris, Petal.Length > 5)
filter(iris, Petal.Length > 5 & Species == "setosa")
filter(iris, Petal.Length > 5, Species == "setosa") the comma is a shorthand for &
filter(iris, !Species == "setosa")
```

D. Pipe Example with MaggriteR (ref: Rene Magritte This is not a pipe)
The long Way, before nesting or multiple variables
```{r long way}
data1 <- filter(iris, Petal.Length > 6)
data2 <- select(data1, Petal.Length, Species)
```

With DPLYR:

```{r pipes}
select(
  filter(iris, Petal.Length > 6),
  Petal.Length, Species)
```

using pipes with the data variable

```{r pipes and verbs}
iris %>%
  filter(iris, Petal.Length > 6) %>%
  select(Petal.Length, Species)
```


using the . to specify where the incoming variable will be piped to:
```{r pipes .}
iris %>%
 myFunction(arg1, arg2 = .)

iris %>%
  filter( ., Species == "setosa")
```

Other magrittr examples

```{r pipes .}
iris %>%
  filter(Petal.Length > 2.0) %>%
  select(1:3)

iris %>%
  select(contains("Width")) %>%
  arrange(Petal.Width) %>%
  head()

iris %>%
  filter(Petal.Width == "versicolor") %>%
  arrange(desc(Sepal.Width))

iris %>%
  filter(Sepal.Width > 1) %>%
  View()

iris %>%
  filter(Petal.Width  == 0.1) %>%
  select(Sepal.Width) %>%
  unique()
```

a second way to get the unique values
```{r pipes .}
iris %>%
  filter(Petal.Width  == 0.1) %>%
  distinct(Sepal.Width)
```


E. Mutate: adds new variables and preserves existing; transmute() drops existing variables
```{r pipes .}
iris %>%
  mutate(highSpecies = Sepal.Width > 6) %>%
  head()


iris %>%
  mutate(size = Sepal.Width + Petal.Width) %>%
  head()


iris %>%
  mutate(MeanPetal.Width = mean(Petal.Width, na.rm = TRUE),
         greaterThanMeanPetal.Width = ifelse(Petal.Width > MeanPetal.Width, 1, 0)) %>%
  head()


iris %>%
  mutate(buckets = cut(Petal.Width, 3)) %>%
  head()


iris %>%
  mutate(Petal.WidthBuckets = case_when(Petal.Width < 50 ~ "Low",
                               Petal.Width >= 60 & Sepal.Width < 5 ~ "Med",
                               Petal.Width >= 70 ~ "High")) %>%
  head()
```

E. Group_by and Summarise: used on grouped data created by group_by().
The output will have one row for each group.
```{r group by}
iris %>%
  summarise(Petal.WidthMean = mean(Petal.Width),
            Petal.WidthSD = sd(Petal.Width))

iris %>%
  group_by(Petal.Width) %>%
  mutate(Petal.WidthMean = mean(Petal.Width))

iris %>%
  group_by(Petal.Width) %>%
  summarise(Petal.WidthMean = mean(Petal.Width))

iris %>%
  group_by(Petal.Width, Species) %>%
  summarise(count = n())
```

F. Slice: Slice does not work with relational databases because they have no intrinsic notion of row order.
If you want to perform the equivalent operation, use filter() and row_number().
```{r slice}
iris %>%
  slice(2:7)
```

Other verbs within DPLYR: 'Scoped verbs'

```{r verb examples}
# ungroup
iris %>%
  group_by(Petal.Width, Species) %>%
  summarise(count = n()) %>%
  ungroup()

# Summarise_all
iris %>%
  select(1:4) %>%
  summarise_all(mean)


iris %>%
  select(1:4) %>%
  summarise_all(funs(mean, min))


iris %>%
  summarise_all(~length(unique(.)))

# summarise_at
iris %>%
  summarise_at(vars(-Petal.Width), mean)


iris %>%
  summarise_at(vars(contains("Petal.Width")), funs(mean, min))

# summarise_if
iris %>%
  summarise_if(is.numeric, Petal.Width)

iris %>%
  summarise_if(is.factor, ~length(unique(.)))

# other verbs:
iris %>%
  mutate_if(is.factor, as.character) %>%
  str()


iris %>%
  mutate_at(vars(contains("Width")), ~ round(.))


iris %>%
  filter_all(any_vars(is.na(.)))


iris %>%
  filter_all(all_vars(is.na(.)))

# Rename
iris %>%
  rename(Species == "Sertosa") %>%
  head()

# And finally: make a test and save
test <- iris %>%
  group_by(Petal.Width) %>%
  summarise(MeanPetal.Width = mean(Petal.Width))
```


