# Demo for dplyr

```{r load libraries for dplyr, message= FALSE}
# Load data and dependencies:
library(dplyr)

data(iris)
```

Explore the iris data

```{r explore data, eval = FALSE}
head(iris)
pairs(iris)
str(iris)
summary(iris)
```

A. **Select**: keeps only the variables you mention

```{r explore, eval = FALSE}
select(iris, 1:3)
select(iris, Petal.Width, Species)
select(iris, contains("Petal.Width"))
select(iris, starts_with("Species"))

```

B. **Arrange**: sort a variable in descending order

```{r arrange, eval = FALSE}
arrange(iris, Sepal.Length)
arrange(iris, desc(Sepal.Length))
arrange(iris, Sepal.Length, desc(Sepal.Width))
```

C. **Filter**: find rows/cases where conditions are true
Note: rows where the condition evaluates to NA are dropped

```{r conditions, eval = FALSE}
filter(iris, Petal.Length > 5)
filter(iris, Petal.Length > 5 & Species == "setosa")
filter(iris, Petal.Length > 5, Species == "setosa") #the comma is a shorthand for &
filter(iris, !Species == "setosa")
```

D. **Pipe Example with MaggriteR** (ref: Rene Magritte This is not a pipe)
The long Way, before nesting or multiple variables
```{r long way, eval = FALSE}
data1 <- filter(iris, Petal.Length > 6)
data2 <- select(data1, Petal.Length, Species)
```

With **DPLYR**:

```{r pipes}
select(
  filter(iris, Petal.Length > 6),
  Petal.Length, Species) %>%
  head()
```

Using pipes with the data variable

```{r pipes and verbs}
iris %>%
  filter(Petal.Length > 6) %>%
  select(Petal.Length, Species) %>%
  head()
```

Using the . to specify where the incoming variable will be piped to: 
- myFunction(arg1, arg2 = .)
 
```{r pipes with filter, eval = FALSE}
iris %>%
  filter(., Species == "versicolor")
```

Other magrittr examples:

```{r unique, eval = FALSE}
iris %>%
  filter(Petal.Length > 2.0) %>%
  select(1:3)

iris %>%
  select(contains("Width")) %>%
  arrange(Petal.Width) %>%
  head()

iris %>%
  filter(Petal.Width == "versicolor") %>%
  arrange(desc(Sepal.Width))

iris %>%
  filter(Sepal.Width > 1) %>%
  View()

iris %>%
  filter(Petal.Width  == 0.1) %>%
  select(Sepal.Width) %>%
  unique()
```

a second way to get the unique values:

```{r distinct }
iris %>%
  filter(Petal.Width  == 0.1) %>%
  distinct(Sepal.Width)
```


E. **Mutate**: adds new variables and preserves existing; transmute() drops existing variables

```{r dplyr verbs, eval = FALSE}
iris %>%
  mutate(highSpecies = Sepal.Width > 6) %>%
  head()


iris %>%
  mutate(size = Sepal.Width + Petal.Width) %>%
  head()


iris %>%
  mutate(MeanPetal.Width = mean(Petal.Width, na.rm = TRUE),
         greaterThanMeanPetal.Width = ifelse(Petal.Width > MeanPetal.Width, 1, 0)) %>%
  head()


iris %>%
  mutate(buckets = cut(Petal.Width, 3)) %>%
  head()


iris %>%
  mutate(Petal.WidthBuckets = case_when(Petal.Width < 1 ~ "Low",
                               Petal.Width >= 2 & Sepal.Width < 3 ~ "Med",
                               Petal.Width >= 4 ~ "High")) %>%
  head()
```

E. **Group_by and Summarise**: used on grouped data created by group_by().
The output will have one row for each group.
```{r group by, eval = FALSE}
iris %>%
  summarise(Petal.WidthMean = mean(Petal.Width),
            Petal.WidthSD = sd(Petal.Width))

iris %>%
  group_by(Petal.Width) %>%
  mutate(Petal.WidthMean = mean(Petal.Width))

iris %>%
  group_by(Petal.Width) %>%
  summarise(Petal.WidthMean = mean(Petal.Width))

iris %>%
  group_by(Petal.Width, Species) %>%
  summarise(count = n())
```

F. **Slice**: Slice does not work with relational databases because they have no intrinsic notion of row order. If you want to perform the equivalent operation, use filter() and row_number().

```{r slice}
iris %>%
  slice(2:4) %>%
  head()
```

Other verbs within DPLYR: **Scoped verbs**

```{r verb examples, eval = FALSE}
# ungroup
iris %>%
  group_by(Petal.Width, Species) %>%
  summarise(count = n()) %>%
  ungroup()

# Summarise_all
iris %>%
  select(1:4) %>%
  summarise_all(mean)


iris %>%
  select(1:4) %>%
  summarise_all(funs(mean, min))


iris %>%
  summarise_all(~length(unique(.)))

# summarise_at
iris %>%
  summarise_at(vars(-Petal.Width), mean)


iris %>%
  summarise_at(vars(contains("Petal.Width")), funs(mean, min))

# summarise_if
iris %>%
  summarise_if(is.numeric, mean)

iris %>%
  summarise_if(is.factor, ~length(unique(.)))

# other verbs:
iris %>%
  mutate_if(is.factor, as.character) %>%
  str()


iris %>%
  mutate_at(vars(contains("Width")), ~ round(.))


iris %>%
  filter_all(any_vars(is.na(.)))


iris %>%
  filter_all(all_vars(is.na(.)))

# Rename
iris %>%
  rename("sp" = "Species") %>%
  head()

# And finally: make a test and save
test <- iris %>%
  group_by(Petal.Width) %>%
  summarise(MeanPetal.Width = mean(Petal.Width))
```


