# Working with dates in R

```{r load lubridate among others, message= FALSE}

library(ggplot2)
library(dplyr)
library(ggridges)
library(hflights)
library(lubridate)
library(hflights)
library(fasttime)
library(microbenchmark)
library(anytime)
library(readr)
```

#check lubridate functions for fun :)
```{r find out today, message= FALSE}

today()
now()
# and check your local  timezone
Sys.timezone()
```


Rule to work with dates according to ISO 8601 standard
format is YYYY-MM-DD 

```{r make up x, message= FALSE}

# The date R 3.0.0 was released
x <- "2013-04-03"

# Examine structure of x
str(x)

# Use as.Date() to interpret x as a date
x_date <- as.Date(x)

# Examine structure of x_date
str(x_date)

# Store April 10 2019 as a Date
april_10_2019 <- as.Date("2019-04-10")

# round dates
r_3_4_1 <- ymd_hms("2016-05-03 07:13:28 UTC")

# Round down to day
floor_date(r_3_4_1, unit = "day")

# Round to nearest 5 minutes
round_date(r_3_4_1, unit = "5 minutes")

# Round up to week
ceiling_date(r_3_4_1, unit = "week")

# Subtract r_3_4_1 rounded down to day
r_3_4_1 - floor_date(r_3_4_1, unit = "day")
```


Setting the timezone:

```{r set timezones, message= FALSE}

# Game2: CAN vs NZL in Edmonton
game2 <- mdy_hm("June 11 2015 19:00")

# Game3: CHN vs NZL in Winnipeg
game3 <- mdy_hm("June 15 2015 18:30")

# Set the timezone to "America/Edmonton"
game2_local <- force_tz(game2, tzone = "America/Edmonton")
game2_local

# Set the timezone to "America/Winnipeg"
game3_local <- force_tz(game3, tzone = "America/Winnipeg")
game3_local

# How long does the team have to rest?
as.period(game2_local %--% game3_local)


# What time is game2_local in NZ?
with_tz(game2_local, tzone = "Pacific/Auckland")

# What time is game2_local in Corvallis, Oregon?
with_tz(game2_local, tzone = "America/Los_Angeles")

# What time is game3_local in NZ?
with_tz(game3_local, tzone = "Pacific/Auckland")
```

Examine DepTime and ArrTime in library(hflights) and others:

```{r load hflights, message= FALSE}
library(hflights)
head(hflights$DepTime, 2)
head(hflights$ArrTime, 2)

# Examine structure of time column
str(hflights$DepTime)
str(hflights$ArrTime)

min(hflights$Year)
max(hflights$Year)

tibble::glimpse(hflights) %>%
   head()
```

Are DepTime and ArrTime the same moments
```{r table of departure time, message= FALSE}
table(hflights$DepTime - hflights$ArrTime) %>%
  head()


# A plot using just time
ggplot(hflights, aes(x = DepTime, y = ArrTime)) +
  geom_line(aes(group = make_date(Year, Month, DayofMonth)), alpha = 0.2)
```

Force datetime to Pacific/Auckland:

```{r change datetime, message= FALSE}

hflights_hourly <- hflights %>%
  mutate(
    Dep = make_date(Year, Month, DayofMonth),
    newDep = force_tz(Dep, tzone = "Pacific/Auckland"))

# check whether times changed
## hflights_hourly$newDep == hflights$DepTime
```

Taking differences of datetimes 
- difftime(time1, time2) takes an argument units which specifies the units for the difference. 
- Your options are "secs", "mins", "hours", "days", or "weeks"

The date landing and moment of step

```{r date landing, message= FALSE}
date_landing <- mdy("July 20, 1969")
moment_step <- mdy_hms("July 20, 1969, 02:56:15", tz = "UTC")

# How many days since the first man on the moon?
difftime(today(), date_landing, units = "days")

# How many seconds since the first man on the moon?
difftime(now(), moment_step, units = "secs")

# another example with three dates
mar_11 <- ymd_hms("2017-03-11 12:00:00", 
                  tz = "America/Los_Angeles")
mar_12 <- ymd_hms("2017-03-12 12:00:00", 
                  tz = "America/Los_Angeles")
mar_13 <- ymd_hms("2017-03-13 12:00:00", 
                  tz = "America/Los_Angeles")

# Difference between mar_13 and mar_12 in seconds
difftime(mar_13, mar_12, units = "secs")

# Difference between mar_12 and mar_11 in seconds
difftime(mar_12, mar_11, units = "secs")
```


Getting datetimes into R
Use as.POSIXct to enter the datetime 
```{r datetime in R, message= FALSE}

as.POSIXct("2010-10-01 12:12:00")

# Use as.POSIXct again but set the timezone to `"America/Los_Angeles"`
as.POSIXct("2010-10-01 12:12:00", tz = "America/Los_Angeles")
```


timespans
Add a period of one week to mon_2pm
```{r time spans, message= FALSE}

mon_2pm <- dmy_hm("27 Aug 2018 14:00")
mon_2pm + weeks(1)

# Add a duration of 81 hours to tue_9am
tue_9am <- dmy_hm("28 Aug 2018 9:00")
tue_9am + hours(81)

# Subtract a period of five years from today()
today() - years(5)

# Subtract a duration of five years from today()
today() - dyears(5)

# Arithmetic with timespans
## Time of North American Eclipse 2017
eclipse_2017 <- ymd_hms("2017-08-21 18:26:40")

# Duration of 29 days, 12 hours, 44 mins and 3 secs
synodic <- ddays(29) + dhours(12) + dminutes(44) + dseconds(3)

# 223 synodic months
saros <- 223*synodic

# Add saros to eclipse_2017
eclipse_2017 + saros
```


Generating sequences of datetimes and arithmetics

Add a period of 8 hours to today

```{r date arithmetic, message= FALSE}

today_8am <- today() + hours(8) 

# Sequence of two weeks from 1 to 26
every_two_weeks <- 1:26 * weeks(2)

# Create datetime for every two weeks for a year
today_8am + every_two_weeks

# A sequence of 1 to 12 periods of 1 month
month_seq <- 1:12 * months(1)

# Add 1 to 12 months to today_8am
today_8am + month_seq 

# Replace + with %m+%
today_8am %m+% month_seq


# %m+% and %m-% are operators not functions. 
## That means you don't need parentheses, just put the operator between the two objects to add or subtract.
# Replace + with %m-%
today_8am %m-% month_seq
```


Intervals
The operator %within% tests if the datetime (or interval) on the left hand side is within the interval of the right hand side. New column for interval from start to end date:

```{r new columns with dates, message= FALSE}

hflights_intervals <- hflights %>% 
  mutate(
    start_date = make_datetime(Year, Month, DayofMonth, DepTime), 
    end_date =  make_datetime(Year, Month, DayofMonth, ArrTime),
    visible = start_date %--% end_date)

# The individual elements 
hflights_intervals$visible[14, ] 

# within
hflights_intervals %>% 
  filter(hflights_intervals$start_date %in% hflights_intervals$end_date) %>%
  select(Year, Month, DayofMonth, ArrTime) %>%
  head()

```


#can create an interval by using the operator %--% with two datetimes. For example ymd("2001-01-01") %--% ymd("2001-12-31") creates an interval for the year of 2001.
Once you have an interval you can find out certain properties like its start, end and length with int_start(), int_end() and int_length() respectively.

Create an interval for flights

```{r int_dates, message= FALSE}

flights <- hflights_intervals %>%
  mutate(ints = start_date %--% end_date) 

# Find the length of flights, and arrange
flights %>%
  mutate(length = length(hflights_intervals$start_date)) %>% 
  arrange(desc(length)) %>%
  head()
``` 

Intervals are the most specific way to represent a span of time since they retain information about the exact start and end moments. 
They can be converted to periods and durations exactly: it's possible to calculate both the exact number of seconds elapsed between the start and end date, as well as the perceived change in clock time.New columns for duration and period


Load the readr package which also has build-in functions for dealing with dates: anytime package:

```{r readr for dates, message= FALSE}

# Various ways of writing Sep 10 2009
sep_10_2009 <- c("September 10 2009", "2009-09-10", "10 Sep 2009", "09-10-2009")

# Use anytime() to parse sep_10_2009
library(anytime)
anytime(sep_10_2009)

# Extract the month of hflights 
month(hflights$Month) %>% table()

# How often is the hour before 12 (noon)?
mean(as.POSIXct(hflights$DepTime, origin="1991-01-01"))


# Use wday() to tabulate release by day of the week
wday(hflights$DepTime) %>% table() %>%
  head()

# Add label = TRUE to make table more readable
wday(hflights$DepTime, label = TRUE) %>% table() %>%
  head()

# Create column wday to hold week days
hflights$wday <- wday(hflights$DepTime, label = TRUE)

```

Parsing dates with `lubridate`

```{r parse with lubridate, message= FALSE}
# Parse x 
x <- "2010 September 20th" 
ymd(x)

# Parse y 
y <- "02.01.2010" 
dmy(y)

# Parse z 
z <- "Sep, 12th 2010 14:00" 
mdy_hm(z)

# Specifying an order with `parse_date_time()`
## Specify an order string to parse x
x <- "Monday June 1st 2010 at 4pm"
parse_date_time(x, orders = "ABdyIp")

# Specify order to include both "mdy" and "dmy"
two_orders <- c("October 7, 2001", "October 13, 2002", "April 13, 2003", 
                "17 April 2005", "23 April 2017")
parse_date_time(two_orders, orders = c("mdy", "dmy"))

# Specify order to include "dOmY", "OmY" and "Y"
short_dates <- c("11 December 1282", "May 1372", "1253")
parse_date_time(short_dates, orders = c("dOmY", "OmY", "Y"))

```



Use make_date() to combine year, month and mday:

```{r parse hrflights, message= FALSE}

hflights_dates  <- hflights  %>% 
  mutate(date = make_date(year = Year, month = Month, day = DayofMonth))

# Plot to check work
ggplot(hflights_dates, aes(x = date, y = DepDelay)) +
  geom_line()
```


If you plot a Date on the axis of a plot, you expect the dates to be in calendar order, 
#and that's exactly what happens with plot() or ggplot().

Set the x axis to the date column:
```{r visualize hflights, message= FALSE}

ggplot(hflights, aes(x = Year, y = DayofMonth)) +
  geom_line(aes(group = 1, color = factor(Month)))
```


Outputting pretty dates and times
Create a stamp based on April 04 2019:
```{r time-stamps, message= FALSE}
D <- ymd("2019-04-04") - days(1:5)
stamp("Created on Sunday, Jan 1, 1999 3:34 pm")(D)
```

More on importing and exporting datetimes
Fast parsing with `fasttime`
The fasttime package provides a single function fastPOSIXct()
#designed to read in datetimes formatted according to ISO 8601. 
#Because it only reads in one format, and doesn't have to guess a format, it is really fast!


The arguments to microbenchmark() are just R expressions that you want to time. 
Make sure you match up the names of these arguments to the parsing functions.


Use fastPOSIXct() to parse dates
```{r posix, message= FALSE}
library(fasttime)
library(microbenchmark)

fastPOSIXct(hflights_dates$date) %>% str()

```