# Demo for Data.table 


Load libraries:

```{r load libraries for data.table, message= FALSE}
# Load data.table
library(data.table)
library(bikeshare14)
library(tidyverse)
```


Create the data.table:

```{r make a table}
X <- data.table(id = c("a", "b", "c"), value = c(0.5, 1.0, 1.5))

print(X)
```


Get number of columns in batrips:

```{r get cols}
batrips <- as.data.table(batrips)
col_number <- ncol(batrips)
col_number
```

Print the first 4 rows:

```{r see rows}
head(batrips, 4)
```

Print the last 4 rows:

```{r see tails}
tail(batrips, 4)
```

Print the structure of batrips:

```{r structure}
str(batrips)
```

Filter third row:

```{r row numbers}
row_3 <- batrips[3]
row_3 %>%
  head(3)
```

Filter rows 1 through 2:

```{r rows }
rows_1_2 <- batrips[1:2]
rows_1_2 %>%
  head(2)
```

Filter the 1st, 6th and 10th rows:

```{r row numbers 1,6,10}
rows_1_6_10 <- batrips[c(1, 6, 10)]
rows_1_6_10 %>%
  head()
```

Select all rows except the first two:

```{r exclude rows}
not_first_two <- batrips[-(1:2)]
not_first_two %>%
  head(2)
```

Select all rows except 1 through 5 and 10 through 15:

```{r exclude more}
exclude_some <- batrips[-c(1:5, 10:15)]
exclude_some %>%
  head(2)
```

Select all rows except the first and last:

```{r all rows}
not_first_last <- batrips[-c(1, .N)] 
# Or
# batrips[-c(1, nrow(batrips))]

not_first_last %>%
  head(2)
```

Filter all rows where start_station is "Market at 10th":

```{r start}
trips_mlk <- batrips[start_station == "Market at 10th"]
trips_mlk %>%
  head(2)
```

Filter all rows where start_station is "MLK Library" AND duration > 1600:

```{r filter is }
trips_mlk_1600 <- batrips[start_station == "MLK Library" & duration > 1600]
trips_mlk_1600 %>%
  head(2)
```

Filter all rows where `subscription_type` is not `"Subscriber"`::

```{r filter is not}
customers <- batrips[subscription_type != "Subscriber"]
customers %>%
  head(2)
```

Filter all rows where start_station is "Ryland Park" AND subscription_type is not "Customer":

```{r filter all}
ryland_park_subscribers <- batrips[start_station == "Ryland Park" & subscription_type != "Customer"]
ryland_park_subscribers %>%
  head(2)
```

Filter all rows where end_station contains "Market":
```{r filter ending with}

any_markets <- batrips[end_station %like% "Market"]
any_markets %>%
  head(2)
```

Filter all rows where trip_id is 588841, 139560, or 139562:
```{r filter combinations}

filter_trip_ids <- batrips[trip_id %in% c(588841, 139560, 139562)]
filter_trip_ids %>%
  head(2)
```

Filter all rows where duration is between [5000, 6000]:
```{r filter between}

duration_5k_6k <- batrips[duration %between% c(5000, 6000)]
duration_5k_6k %>%
  head(2)
```

Filter all rows with specific start stations:
```{r filter in}

two_stations <- batrips[start_station %chin% c("San Francisco City Hall", "Embarcadero at Sansome")]
two_stations %>%
  head(2)
```


Selecting columns from a data.table
Select bike_id and trip_id using a character vector:
```{r select some columns}

df_way <- batrips[, c("bike_id", "trip_id")]
df_way %>%
  head(2)
```

Select start_station and end_station cols without a character vector:
```{r start and end with }

dt_way <- batrips[, .(start_station, end_station)]
dt_way %>%
  head(2)
```

Deselect start_terminal and end_terminal columns:
```{r select not }

drop_terminal_cols <- batrips[, !c("start_terminal", "end_terminal")]
drop_terminal_cols %>%
  head(2)
```

Calculate median duration using the j argument:
```{r calculate median}

median_duration <- batrips[, median(duration)]
median_duration %>%
  head()
```

Get median duration after filtering:
```{r median after filtering}

median_duration_filter <- batrips[end_station == "Market at 10th" & subscription_type == "Subscriber", median(duration)]
median_duration_filter %>%
  head()
```

Compute duration of all trips:
```{r duration}

trip_duration <- batrips[, difftime(end_date, start_date, units = "min")]
head(trip_duration) %>%
  head(2)
```

Have the column mean_durn:
```{r make a mean_durn}

mean_duration <- batrips[, .(mean_durn = mean(duration))]
mean_duration %>%
  head(2)
```

Get the min and max duration values:
```{r make a min/max}

min_max_duration <- batrips[, .(min(duration), max(duration))]
min_max_duration %>%
  head(2)
```

Calculate the number of unique values:
```{r calculate unique}

other_stats <- batrips[, .(mean_duration = mean(duration), 
                           last_ride = max(end_date))]
other_stats %>%
  head(2)
```

```{r select min / max}
duration_stats <- batrips[start_station == "Townsend at 7th" & duration < 500, 
                          .(min_dur = min(duration), 
                            max_dur = max(duration))]
duration_stats
```

Plot the histogram of duration based on conditions:
```{r make histogram}

batrips[start_station == "Townsend at 7th" & duration < 500, hist(duration)]
```


Computations by groups
Compute the mean duration for every start_station:
```{r compute by group}

mean_start_stn <- batrips[, .(mean_duration = mean(duration)), by = start_station]
mean_start_stn %>%
  head(2)
```

Compute the mean duration for every start and end station:
```{r compute for start/end}

mean_station <- batrips[, .(mean_duration = mean(duration)), by = .(start_station, end_station)]
mean_station %>%
  head(2)
```

Compute the mean duration grouped by start_station and month:
```{r grouped}

mean_start_station <- batrips[, .(mean_duration = mean(duration)), by = .(start_station, month(start_date))]
mean_start_station %>%
  head(2)
```

Compute mean of duration and total trips grouped by start and end stations:
```{r duration and total}

aggregate_mean_trips <- batrips[, .(mean_duration = mean(duration), 
                                    total_trips = .N), 
                                by = .(start_station, end_station)]
aggregate_mean_trips %>%
  head(2)
```

Compute min and max duration grouped by start station, end station, and month:
```{r min and max}

aggregate_min_max <- batrips[, .(min_duration = min(duration), 
                                 max_duration = max(duration)), 
                             by = .(start_station, end_station, 
                                    month(start_date))]
aggregate_min_max %>%
  head(2)
```

Chaining data.table expressions:
Compute the total trips grouped by start_station and end_station

```{r Compute the total}

trips_dec <- batrips[, .N, by = .(start_station, 
                                  end_station)]
trips_dec %>%
  head(2)
```

Arrange the total trips grouped by start_station and end_station in decreasing order:
```{r Arrange}

trips_dec <- batrips[, .N, by = .(start_station, 
                                  end_station)][order(-N)]
trips_dec %>%
  head(2)
``` 

Top five most popular destinations:
```{r Top five}

top_5 <- batrips[, .N, by = end_station][order(-N)][1:5]
top_5
```

Compute most popular end station for every start station:
```{r Compute most popular}

popular_end_station <- trips_dec[, .(end_station = end_station[1]), 
                                 by = start_station]
popular_end_station %>%
  head(2)
```

Find the first and last ride for each start_station:
```{r  first and last}

first_last <- batrips[order(start_date), 
                      .(start_date = start_date[c(1, .N)]), 
                      by = start_station]
first_last
```

Using .SD (I)

```{r standard deviation}

relevant_cols <- c("start_station", "end_station", 
                   "start_date", "end_date", "duration")
```

Find the row corresponding to the shortest trip per month:

```{r date selection}

shortest <- batrips[, .SD[which.min(duration)], 
                    by = month(start_date), 
                    .SDcols = relevant_cols]
shortest %>%
  head(2)
```

Using .SD (II)
Find the total number of unique start stations and zip codes per month:

```{r date by month}

unique_station_month <- batrips[, lapply(.SD, uniqueN), 
                                by = month(start_date), 
                                .SDcols = c("start_station", "zip_code")]
unique_station_month %>%
  head(2)
```


Adding and updating columns by reference
Add a new column, duration_hour:

```{r duration calculations}
batrips[, duration_hour := duration / 3600]
```

Fix/edit spelling in the second row of start_station:

```{r select based on speeling}
batrips[2, start_station := "San Francisco City Hall 2"]
```

Replace negative duration values with NA:

```{r less or more}
batrips[duration < 0, duration := NA]
```

Add a new column equal to total trips for every start station:

```{r make a new column}
batrips[, trips_N := .N, by = start_station]
```

Add new column for every start_station and end_station:

```{r new column with calculation}
batrips[, duration_mean := mean(duration), by = .(start_station, end_station)]
```

Calculate the mean duration for each month:

```{r calculate mean}
batrips[, mean_dur := mean(duration, na.rm = TRUE), 
            by = month(start_date)]
```

Replace NA values in duration with the mean value of duration for that month:

```{r mean value per month}
batrips[, mean_dur := mean(duration, na.rm = TRUE), 
            by = month(start_date)][is.na(duration), 
                                    duration := mean_dur]
```

Delete the mean_dur column by reference:

```{r mean duration}
batrips[, mean_dur := mean(duration, na.rm = TRUE), 
            by = month(start_date)][is.na(duration), 
                                    duration := mean_dur][, mean_dur := NULL]
```

Add columns using the LHS := RHS form
LHS := RHS form. In the LHS, you specify column names as a character vector and in the RHS, you specify values/expressions to be added inside list() (or the alias, .()):

```{r left and right hand side }
batrips[, c("mean_duration", 
            "median_duration") := .(mean(duration), median(duration)), 
        by = start_station]
```

Add columns using the functional form:

```{r functions}
batrips[, `:=`(mean_duration = mean(duration), 
               median_duration = median(duration)), 
        by = start_station]
```

Add the mean_duration column:

```{r function duration selection}
batrips[duration > 600, mean_duration := mean(duration), 
        by = .(start_station, end_station)]
```

Use read.csv() to import batrips
Fread is much faster!

- system.time(read.csv("batrips.csv"))
- system.time(fread("batrips.csv"))


Import using read.csv():

```{r read in a file}
csv_file <- read.csv("data/sample.csv", fill = NA, quote = "", 
                     stringsAsFactors = FALSE, strip.white = TRUE, 
                     header = TRUE)
csv_file %>%
  head(2)
```

Import using fread():

```{r read in a csv file}
csv_file <- fread("data/sample.csv")
csv_file %>%
  head(2)
```

Check the class of Sex column:

```{r check class}
class(csv_file$Sex)
```

Import using read.csv with defaults:

```{r check structure}
str(csv_file)
```

Select "id" and "val" columns:

```{r read and select some columns}
select_columns <- fread("data/sample.csv", select = c("GEO", "Sex"))
select_columns %>%
  head(2)
```

Drop the "val" column:

```{r read and drop columns}
drop_column <- fread("data/sample.csv", drop = "Sex")
drop_column %>%
  head(2)
```

Import the file while avoiding the warning:

```{r read first three rows}
only_data <- fread("data/sample.csv", nrows = 3)
only_data
```

Import only the metadata:

```{r read while skipping some lines}
only_metadata <- fread("data/sample.csv", skip = 7)
only_metadata %>%
  head(2)
```

Import using read.csv:

```{r read a csv in R}
base_r <- read.csv("data/sample.csv", 
                   colClasses = c(rep("factor", 4), 
                                  "character", 
                                  "numeric"))
str(base_r)
```

Import using fread:

```{r read csv with column classes}
import_fread <- fread("data/sample.csv", 
                      colClasses = list(factor = 1:4, numeric = 7:10))
str(import_fread)
```

Import the file correctly,  use the fill argument to ensure all rows are imported correctly:

```{r read with fill}
correct <- fread("data/sample.csv", fill = TRUE)
correct %>%
  head(2)
```

Import the file using na.strings
The missing values are encoded as "##". Note that fread() handles an empty field ,, by default as NA

```{r read csv}
missing_values <- fread("data/sample.csv", na.strings = "##") 
missing_values %>%
  head(2)
```

Write dt to fwrite.txt:
- fwrite(dt, "fwrite.txt")

Import the file using readLines():

```{r readlines into R}
readLines("data/sample.csv") %>%
  head(2)
```

Write batrips_dates to file using "ISO" format:
- fwrite(batrips_dates, "iso.txt", dateTimeAs = "ISO")

Write batrips_dates to file using "squash" format:
- fwrite(batrips_dates, "squash.txt", dateTimeAs = "squash")


