# Tests for experiments

Prior to performing experiments, we need to set the dependent variables (outcome), and independent variables (explanatory variables).

Other experimental components to consider include randomization, replication, blocking

```{r load libraries ggplot broom, message= FALSE}
# load dependencies
library(ggplot2) 
library(broom)
library(tidyverse)
library(pwr)
library(haven)
library(simputation)
library(sampling)
library(agricolae)
library(naniar)
library(DescTools)
library(mice)
```


load data: Dataset is on the Effect of Vitamin C on Tooth Growth in Guinea Pigs:
```{r load data, message= FALSE}
data(ToothGrowth) 

ToothGrowth %>%
  head(2)
```

Perform a two-sided t-test:
```{r perform t-test on one variable, message= FALSE}

t.test(x = ToothGrowth$len, alternative = "two.sided", mu = 18)
```

Perform a t-test
```{r perform t-test on two variables, message= FALSE}

ToothGrowth_ttest <- t.test(len ~ supp, data = ToothGrowth)
ToothGrowth_ttest
```


Tidy ToothGrowth_ttest:
```{r tidy visualize, message= FALSE}
tidy(ToothGrowth_ttest)
```


Replication:
Count number of observations for each combination of supp and dose
```{r tidy, message= FALSE}

ToothGrowth %>% 
  count(supp, dose) 
```

Blocking:
Create a boxplot with geom_boxplot()
aov() creates a linear regression model by calling lm() and examining results with anova() all in one function call.

```{r Visualize, message= FALSE}

ggplot(ToothGrowth, aes(x = dose, y = len)) + 
  geom_boxplot()
```


Create ToothGrowth_aov and 
Examine ToothGrowth_aov with summary():

```{r aov, message= FALSE}

ToothGrowth_aov <- aov(len ~ dose + supp, data = ToothGrowth)
summary(ToothGrowth_aov)
```


Hypothesis Testing (null and alternative) with `pwr` package
one sided and two sided tests: 
- type ?t.test to find out more
```{r pwr less, message= FALSE}

#Less than
t.test(x = ToothGrowth$len,
       alternative = "less",
       mu = 18)
```

```{r pwr greater, message= FALSE}

# Greater than
t.test(x = ToothGrowth$len,
       alternative = "greater",
       mu = 18)
```


It turns out the mean of len is actually very close to 18, so neither of these tests tells us much about the mean of tooth length.
?pwr.t.test()

Calculate sample size:

```{r pwr sample size, message= FALSE}

pwr.t.test(n = NULL,
           d = 0.25,  # small effect size of 0.25
           sig.level = 0.05, 
           type = "one.sample", 
           alternative = "greater", 
           power = 0.8)
```


Calculate power:

```{r pwr power, message= FALSE}

pwr.t.test(n = 100,
           d = 0.35,
           sig.level = 0.1,
           type = "two.sample",
           alternative = "two.sided",
           power = NULL)
```

power for multiple groups:

```{r pwr power for multiple groups k, message= FALSE}

pwr.anova.test(k = 3,
               n = 20,
               f = 0.2, #effect size
               sig.level = 0.05,
               power = NULL)
```


Anova tests (for multiple groups) can be done in two ways

Basic Experiments for exploratory data analysis including A/B testing

get data:

```{r get data txhousing, message= FALSE}

data(txhousing)

txhousing %>%
  head(2)
```


remove NAs:
```{r omit NA, message= FALSE}
tx_housing <- na.omit(txhousing)

# Examine the variables with glimpse()
glimpse(tx_housing)
```

Find median and means with summarize():

```{r find med/mean, message= FALSE}
tx_housing %>% 
  summarize(median(volume), mean(sales), mean(inventory))

```

Use ggplot2 to build a bar chart of purpose:

```{r visualize txhousing, message= FALSE}
ggplot(data=tx_housing, aes(x = city)) + 
  geom_bar() +
  coord_flip()
```
Use recode() to create the new purpose_recode variable

```{r recode, message= FALSE}
tx_housing$city_recode <- tx_housing$city %>%
  recode("Bay Area" = "California",
         "El Paso" = "California")
```

Build a linear regression model, purpose_recode_model:
```{r recode lm, message= FALSE}
purpose_recode_model <- lm(sales ~ city_recode, data = tx_housing)

# Examine results of purpose_recode_model
summary(purpose_recode_model)
```



Get anova results and save as purpose_recode_anova:

```{r recode anova, message= FALSE}

purpose_recode_anova <- anova(purpose_recode_model)

# Print purpose_recode_anova
purpose_recode_anova

```


Examine class of purpose_recode_anova:
```{r recode class, message= FALSE}

class(purpose_recode_anova)
```

Use aov() to build purpose_aov:
```{r recode aov, message= FALSE}

# Analysis of variance
purpose_aov <- aov(sales ~ city_recode, data = tx_housing)
```

Conduct Tukey's HSD test to create tukey_output:
```{r recode tukey test, message= FALSE}

tukey_output <- TukeyHSD(purpose_aov, "city_recode", conf.level = 0.95)

# Tidy tukey_output to make sense of the results
tidy(tukey_output)
```


Multiple factor experiments:
Use aov() to build purpose_emp_aov
```{r manova, message= FALSE}

purpose_emp_aov <- aov(sales ~ city_recode + volume , data = tx_housing)

# Print purpose_emp_aov to the console
# purpose_emp_aov

#Call summary() to see the p-values:
summary(purpose_emp_aov)
```



Model validation
Pre-modeling exploratory data analysis
Examine the summary of sales

```{r exploratory, message= FALSE}
summary(tx_housing$sales)
```

Examine sales by volume:

```{r examine sales, message= FALSE}
tx_housing %>% 
  group_by(volume) %>% 
  summarize(mean = mean(sales), var = var(sales), median = median(sales))
```

Make a boxplot of sales by volume

```{r box plot of sales a/b test, message= FALSE}
ggplot(tx_housing, aes(x = volume, y = sales)) + 
  geom_boxplot()
```

Use aov() to create volume_aov plus call summary() to print results
```{r aov and summary, message= FALSE}
volume_aov <- aov(volume ~ sales, data = tx_housing)
summary(volume_aov)
```

Post-modeling validation plots + variance
For a 2x2 grid of plots:

```{r qq plots, message= FALSE}
par(mfrow = c(2, 2))

# Plot grade_aov
plot(volume_aov)
```

Bartlett's test for homogeneity of variance
We can test for homogeneity of variances using bartlett.test(), which takes a formula and a dataset as inputs:
- bartlett.test(volume ~ sales, data = tx_housing)

Conduct the Kruskal-Wallis rank sum test:
kruskal.test() to examine whether volume varies by sales when a non-parametric model is employed

```{r Kruskal-Wallis, message= FALSE}
kruskal.test(volume ~ sales,
             data = tx_housing)
```

The low p-value indicates that based on this test, we can be confident in our result, which we found across this experiment, that volume varies by sales


Sampling [randomized experiments]

load data from NHANES dataset
https://wwwn.cdc.gov/nchs/nhanes/continuousnhanes/default.aspx?BeginYear=2015

Import the three datasets using read_xpt():

```{r import data from urls, message= FALSE}
nhanes_demo <- read_xpt(url("https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/DEMO_I.XPT"))
nhanes_bodymeasures <- read_xpt(url("https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/BMX_I.XPT"))
nhanes_medical <- read_xpt(url("https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/MCQ_I.XPT"))
```

Merge the 3 datasets you just created to create nhanes_combined:

```{r merge data, message= FALSE}

nhanes_combined <- list(nhanes_demo, nhanes_medical, nhanes_bodymeasures) %>%
  Reduce(function(df1, df2) inner_join(df1, df2, by = "SEQN"), .)
```

Fill in the dplyr code:
```{r summarize merged data, message= FALSE}

nhanes_combined %>% 
  group_by(MCQ035) %>% 
  summarize(mean = mean(INDHHIN2, na.rm = TRUE))
```

Fill in the ggplot2 code:

```{r  plot data, message= FALSE}
nhanes_combined %>% 
  ggplot(aes(as.factor(MCQ035), INDHHIN2)) +
  geom_boxplot() +
  labs(x = "Disease type",
       y = "Income")
```

NHANES Data Cleaning
Filter to keep only those greater than 16:
```{r  filter age, message= FALSE}
nhanes_filter <- nhanes_combined %>% filter(RIDAGEYR > 16)
```

Load simputation & impute bmxwt by riagendr: library(simputation)
```{r use simputation, message= FALSE}

nhanes_final <- simputation::impute_median(nhanes_filter, INDHHIN2 ~ RIDAGEYR)

```

Recode mcq365d with recode() & examine with count():
```{r  recode mcq365d, message= FALSE}

nhanes_final$mcq365d <- recode(nhanes_final$MCQ035, 
                               `1` = 1,
                               `2` = 2,
                               `9` = 2)
nhanes_final %>% count(MCQ035)
```

Resampling NHANES data:
Use sample_n() to create nhanes_srs:
```{r  resample, message= FALSE}

nhanes_srs <- nhanes_final %>% sample_n(2500)
```

Create nhanes_stratified with group_by() and sample_n()
```{r  sample_n and group_by, message= FALSE}

nhanes_stratified <- nhanes_final %>% group_by(RIDAGEYR) %>% sample_n(2000, replace = TRUE)

nhanes_stratified %>% 
  count(RIDAGEYR)
```


Load sampling package and create nhanes_cluster with cluster(): library(sampling)

```{r  sample_n, message= FALSE}

nhanes_cluster <- cluster(nhanes_final, c("INDHHIN2"), 6, method = "srswor")

```

Randomized complete block designs (RCBD): use library(agricolae)
block = experimental groups are blocked to be similar (e.g. by sex)
complete = each treatment is used the same of times in every block
randomized = the treatment is assigned randomly inside each block

Create designs using ls():
```{r  agricolae library, message= FALSE}

designs <- ls("package:agricolae", pattern = "design")
print(designs)
```

Use str() to view design.rcbd's criteria:
```{r structure of design, message= FALSE}

str(design.rcbd)
```

Build treats and rep

```{r build letters, message= FALSE}

treats <- LETTERS[1:5]
blocks <- 4
blocks
```


NHANES RCBD:
Build my_design_rcbd and view the sketch
```{r build and view, message= FALSE}

my_design_rcbd <- design.rcbd(treats, r = blocks, seed = 42)
my_design_rcbd$sketch
```

Use aov() to create nhanes_rcbd:
```{r aov for nhanes, message= FALSE}

nhanes_rcbd <- aov(INDHHIN2 ~ MCQ035 + RIDAGEYR, data = nhanes_final)
```
Check results of nhanes_rcbd with summary():
```{r result summary, message= FALSE}

summary(nhanes_rcbd)
```
Print mean weights by mcq365d and riagendr:
```{r mean weights, message= FALSE}

nhanes_final %>% 
  group_by(MCQ035, RIDAGEYR) %>% 
  summarize(mean_ind = mean(INDHHIN2, na.rm = TRUE))
```

RCBD Model Validation
Set up the 2x2 plotting grid and plot nhanes_rcbd
```{r 2x2 grid, message= FALSE}

par(mfrow = c(2, 2))
plot(nhanes_rcbd)
```

Run the code to view the interaction plots:
```{r view final, message= FALSE}

with(nhanes_final, interaction.plot(MCQ035, RIDAGEYR, INDHHIN2))
```


Balanced incomplete block design (BIBD)
Balanced = each pair of treatment occur together in a block an equal of times
Incomplete = not every treatment will appear in every block

Use str() to view design.bibd's criteria
str(design.bib) 

Columns are a blocking factor

create my_design_bibd_1
```{r design.bib, message= FALSE}

my_design_bibd_1 <- agricolae::design.bib(LETTERS[1:3], k = 3, seed = 42)
```

create my_design_bibd_2
```{r create with k of 3, message= FALSE}

my_design_bibd_2 <- design.bib(LETTERS[1:8], k = 8, seed = 42)
```



create my_design_bibd_3:
  
```{r create with k of 4, message= FALSE}

my_design_bibd_3 <- design.bib(LETTERS[1:4], k = 4, seed = 42)
my_design_bibd_3$sketch
```


Build the data.frame:
```{r build data, message= FALSE}

creatinine <- c(1.98, 1.97, 2.35, 2.09, 1.87, 1.95, 2.08, 2.01, 1.84, 2.06, 1.97, 2.22)
food <- as.factor(c("A", "C", "D", "A", "B", "C", "B", "C", "D", "A", "B", "D"))
color <- as.factor(rep(c("Black", "White", "Orange", "Spotted"), each = 3))
cat_experiment <- as.data.frame(cbind(creatinine, food, color))
```

Create cat_model and examine with summary():

```{r create model, message= FALSE}

cat_model <- aov(creatinine ~ food + color, data = cat_experiment)
summary(cat_model)
```

Calculate lambda, where lamdba is a measure of proportional reduction in error in cross tabulation analysis:

```{r calculate lambda, message= FALSE}
DescTools::Lambda(cat_experiment, direction = c("symmetric", "row", "column"), conf.level = NA)
```

Create weightlift_model & examine results:
```{r create model /results, message= FALSE}

weightlift_model <- aov(MCQ035 ~ INDHHIN2 + RIDAGEYR, data = nhanes_final)
summary(weightlift_model)
```


Latin Squares Design
Key assumption: the treatment and two blocking factors do NOT interact
Two blocking factors (instead of one)
Analyze like RCBD

Mean, var, and median of Math score by Borough:

```{r calculate mean, var, median, message= FALSE}

sat_scores <- read.csv(url("https://data.ct.gov/api/views/kbxi-4ia7/rows.csv?accessType=DOWNLOAD"))

sat_scores %>%
  group_by(District, Test.takers..2012) %>% 
  summarize(mean = mean(Test.takers..2012, na.rm = TRUE),
            var = var(Test.takers..2012, na.rm = TRUE),
            median = median(Test.takers..2012, na.rm = TRUE)) %>%
  head()
```

Dealing with Missing Test Scores
Examine missingness with miss_var_summary() and library(mice):
```{r Examine missingness with md.pattern(), message= FALSE}

sat_scores %>% miss_var_summary()
sat_scores <- na.omit(sat_scores)

mice::md.pattern(sat_scores)
```


Impute the Math score by Borough:
```{r impute median, message= FALSE}

sat_scores_2 <- simputation::impute_median(sat_scores, Test.takers..2012 ~ District)
#Convert Math score to numeric
sat_scores$Average_testtakers2012 <- as.numeric(sat_scores$Test.takers..2012)
```

Examine scores by Borough in both datasets, before and after imputation:
```{r cexamine scores, message= FALSE}

sat_scores %>% 
  group_by(District) %>% 
  summarize(median = median(Test.takers..2012, na.rm = TRUE), 
            mean = mean(Test.takers..2012, na.rm = TRUE))
sat_scores_2 %>% 
  group_by(District) %>% 
  summarize(median = median(Test.takers..2012), 
            mean = mean(Test.takers..2012))
```

Drawing Latin Squares with agricolae

Design a LS with 5 treatments A:E then look at the sketch
```{r ls with 5 treatments, message= FALSE}
my_design_lsd <- design.lsd(trt = LETTERS[1:5], seed = 42)
my_design_lsd$sketch
```

Build nyc_scores_ls_lm:

```{r ls_lm, message= FALSE}
sat_scores_ls_lm <- lm(Test.takers..2012 ~ Test.takers..2013 + District,
                       data = sat_scores)

# Tidy the results with broom
tidy(sat_scores_ls_lm) %>%
  head()
```

Examine the results with anova:

```{r anova for scores, message= FALSE}
anova(sat_scores_ls_lm)
```


Graeco-Latin Squares
three blocking factors (when there is treatments)
Key assumption: the treatment and two blocking factors do NOT interact
Analyze like RCBD

Drawing Graeco-Latin Squares with agricolae

Create trt1 and trt2
Create my_graeco_design
```{r graeco design, message= FALSE}
trt1 <- LETTERS[1:5]
trt2 <- 1:5
my_graeco_design <- design.graeco(trt1, trt2, seed = 42)
```

Examine the parameters and sketch:

```{r parameters/sketch, message= FALSE}
my_graeco_design$parameters
my_graeco_design$sketch
```

Create a boxplot of scores by District, with a title and x/y axis labels:
```{r scores and titles, message= FALSE}
ggplot(sat_scores, aes(District, Test.takers..2012)) +
  geom_boxplot() + 
  labs(title = "Average SAT Math Scores by District in 2012",
       x = "District",
       y = "Test Takers in 2012")
```

Build sat_scores_gls_lm:

```{r gls_lm, message= FALSE}
sat_scores_gls_lm <- lm(Test.takers..2012 ~ Test.takers..2013 + District + School,
                        data = sat_scores)

# Tidy the results with broom
tidy(sat_scores_gls_lm) %>%
  head()
```

Examine the results with anova
```{r anova examination, message= FALSE}
anova(sat_scores_gls_lm)
```


Factorial Experiment Design
2 or more factor variables are combined and crossed
All of the possible interactions between factors are considered as effect on outcome
e.g. high/low water on high/low light


Build the boxplot for the district vs. test taker score:

```{r boxplot with scores, message= FALSE}
ggplot(sat_scores,
       aes(District, Test.takers..2012)) + 
  geom_boxplot()
```

Create sat_scores_factorial and examine the results:
```{r aov and tidy, message= FALSE}
sat_scores_factorial <- aov(Test.takers..2012 ~ Test.takers..2013 * District * School, data = sat_scores)

tidy(sat_scores_factorial) %>%
  head()
```

Evaluating the sat_scores Factorial Model

Use shapiro.test() to test the outcome:
```{r shapiro test, message= FALSE}
shapiro.test(sat_scores$Test.takers..2013)
```
