
# Demo for A/B testing 

```{r dependencies, message= FALSE}
# load dependencies
library(tidyverse)
library(powerMediation)
library(broom)
library(pwr)
library(gsDesign)
library(powerMediation)
```

Read in data:

```{r read-in the data, message= FALSE}

fileLocation <- "http://stat.columbia.edu/~rachel/datasets/nyt1.csv"
click_data <- read.csv(url(fileLocation))
```

Find oldest and most recent age:

```{r check max/min, message= FALSE}

min(click_data$Age)
max(click_data$Age) 
```


Compute baseline conversion rates:

```{r baseline rates, message= FALSE}

click_data %>%
  summarize(impression_rate = mean(Impressions))
```

determine baseline for genders:

```{r baseline genders, message= FALSE}

click_data %>%
  group_by(Gender) %>%
  summarize(impression_rate = mean(Impressions))
```


determine baseline for clicks:

```{r baseline clicks, message= FALSE}

click_data_age<- click_data %>%
  group_by(Clicks, Age) %>%
  summarize(impression_rate = mean(Impressions))
```

visualize baselines:

```{r visualize baselines, message= FALSE}

ggplot(click_data_age, aes(x = Age, y = impression_rate)) +
  geom_point() +
  geom_line() 
```


Experimental design, power analysis, and t-tests

run power analysis:
learn more here: help(SSizeLogisticBin)

```{r sample size for clicks, message= FALSE}

total_sample_size <- SSizeLogisticBin(p1 = 0.2, # conversion rate for control condition
                                      p2 = 0.3, # conversion rate for expected conversion rate for test condition: backed by previous data (e.g.30% conversion rate to get 10% boost)
                                      B = 0.5, # most commonly used
                                      alpha = 0.05, # most commonly used
                                      power = 0.8) # most commonly used
total_sample_size
total_sample_size /2 # per condition
```

can use a ttest or linear regression for statistical tests:
lm is used when more variables are in data but similar to t-test
```{r lm for clicks, message= FALSE}

lm(Gender ~ Clicks, data = click_data) %>%
  summary()

# t.test(Gender ~ Clicks, data = click_data) %>%
#  summary()
```

Analyzing results
Group and summarize
```{r click summary, message= FALSE}

click_data_groups <- click_data %>%
  group_by(Clicks, Age) %>%
  summarize(impression_rate = mean(Impressions))
```

Make plot of conversion rates for clicks:
```{r visualize clicks, message= FALSE}

ggplot(click_data_groups,
       aes(x = Age,
           y = impression_rate,
           color = Clicks,
           group = Clicks)) +
  geom_point(size = 3) +
  geom_line(lwd = 1)
```

Make plot of conversion rates for clicks 
(can add intercepts and interaction of two variables):
```{r plot clicks, message= FALSE}

ggplot(click_data_groups,
       aes(x = Age,
           y = impression_rate,
           color = Clicks,
           group = interaction(Clicks, impression_rate))) +
  geom_point(size = 3) +
  geom_line(lwd = 1) +
  geom_vline(xintercept = as.numeric(as.Date("2018-02-15"))) 

```

Check for glm documentation
family can be used to express different error distributions.
?glm

Run logistic regression to analyze model outputs:
```{r run glm for clicks, message= FALSE}

experiment_results <- glm(Gender ~ Clicks,
                          family = "binomial",
                          data = click_data) %>%
  tidy()

experiment_results
```


Follow-up experimentations to test assumptions

Designing follow-up experiments since A/B testing is a continuous loops
i.e. make new dataframes and compute various other conversion rate differences
can use spread() to reformat data

```{r test experiment for click data, message= FALSE}

click_data_new_groups <- click_data %>%
  group_by(Clicks, Age) %>%
  summarize(impression_rate = mean(Impressions)) %>% 
  spread(Clicks, impression_rate)
```
Compute summary statistics:

```{r compute summary for clicks, message= FALSE}

mean(click_data_new_groups$Age, na.rm = TRUE)
sd(click_data_new_groups$Age, na.rm = TRUE)
```

Run logistic regression and power analysis
Run power analysis for logistic regression

```{r SS & power, message= FALSE}

total_sample_size <- SSizeLogisticBin(p1 = 0.49,
                                      p2 = 0.64,
                                      B = 0.5,
                                      alpha = 0.05,
                                      power = 0.8)
total_sample_size
```

View summary of data:
```{r means of click impressions, message= FALSE}

new_data <- click_data %>%
  group_by(Clicks) %>%
  summarize(impression_rate = mean(Impressions)/10)

# Run logistic regression to analyze model outputs
followup_experiment_sep_results <- glm(impression_rate ~ Clicks,
                                       family = "binomial",
                                       data = new_data) %>%
  tidy()

followup_experiment_sep_results
```


Specifics of A/B Testing= use of experimental design to compare 2 or more variants of a design
Test Types: A/B, A/A, A/B/N test conditions 
Assumptions to test: within group vs. between group experiments

e.g. plotting A/A data
Compute conversion rates for A/A experiment:

```{r compute conversion rates, message= FALSE}

click_data_sum <- click_data %>%
  group_by(Signed_In) %>%
  summarize(impression_rate = mean(Impressions)/10)
click_data_sum
```

Plot conversion rates for two conditions:
```{r plot conversion rates, message= FALSE}

ggplot(click_data_sum,
       aes(x = Signed_In, y = impression_rate)) +
  geom_bar(stat = "identity")  
  
#Based on these bar plots the two A conditions look very similar. That's good!
```


Run logistic regression to analyze model outputs:

```{r use broom to run glm, message= FALSE}

aa_experiment_results <- glm(Signed_In ~ impression_rate,
                             family = "binomial",
                             data = click_data_sum) %>%
  tidy()
aa_experiment_results
```


Confounding variables: element that can affect the truth of A/B exp
change one element at a time to know the change you are testing
Need to also consider the side effects 
procedures are the same as above

Power analysis requires 3 variables: power (1-beta) , significance level (alpha or p-value), effect size
as power goes up, so does the of data points needed
as significance level goes up (i.e. more significant), so do of data points needed
as effect sizw increase, of data points decrease
ttest (linear regression) can be used for continuous dependent variable (e.g. time spent on a website)

```{r use pwr for lm, message= FALSE}

pwr.t.test(power = 0.8,
           sig.level = 0.05,
           d = 0.6)  # d = effect size 

pwr.t.test(power = 0.8,
           sig.level = 0.05,
           d = 0.2) #(see more on experimental design)

```

Load package to run power analysis: library(powerMediation)

logistic regression can be used for categorical dependent variable (e.g. click or not click)
Run power analysis for logistic regression

```{r use powerMediation package, message= FALSE}

total_sample_size <- SSizeLogisticBin(p1 = 0.17, # assuming a control value of 17%
                                      p2 = 0.27, # assuming 10% increase in the test condition
                                      B = 0.5,
                                      alpha = 0.05,
                                      power = 0.8)
total_sample_size
```


Stopping rules and sequential analysis
procedures that allow interim analyses in pre-defined points = sequential analysis

```{r use powerMediation with gsdesign, message= FALSE}

seq_analysis <- gsDesign(k=4, # number of times you want to look at the data
                         test.type = 1,
                         alpha = 0.05,
                         beta = 0.2, # power = 1-beta so power is 0.8
                         sfu = "Pocock") # spending function to figure out how to update p-values
seq_analysis

```

```{r random generation of sequences, message= FALSE}

max_n <- 1000
max_n_per_group <- max_n / 2
stopping_points <- max_n_per_group * seq_analysis$timing
stopping_points
```


Run sequential analysis:

```{r use gsdesign for power analysis, message= FALSE}

seq_analysis_3looks <- gsDesign(k = 3,
                                test.type = 1,
                                alpha = 0.05,
                                beta = 0.2,
                                sfu = "Pocock")
seq_analysis_3looks
```

Fill in max number of points and compute points per group and find stopping points

```{r random generation of sequences max.min, message= FALSE}

max_n <- 3000
max_n_per_group <- max_n / 2
stopping_points = max_n_per_group * seq_analysis_3looks$timing
stopping_points

```

Multivariate testing (i.e. more than one independent variable in the experiment)

Compute summary values for four conditions
```{r use broom to visualize clicks, message= FALSE}

new_click_data <- click_data %>%
  group_by(Age, Gender, Clicks) %>%
  summarize(impression_mean = mean(Impressions))

# Plot summary values for four conditions
ggplot(new_click_data,
       aes(x = Gender,
           y = impression_mean,
           color = Clicks,
           fill = Age)) +
  geom_bar(stat = "identity", position = "dodge")
```

```{r broom with lm, message= FALSE}

multivar_results <- lm(Age ~ Gender * Clicks, data = click_data) %>%
  tidy()

multivar_results$p.value #none are significant
multivar_results
```

Organize variables and run logistic regression:

```{r run lm with variable, message= FALSE}

new_click_data_results <- click_data %>%
  mutate(gender = factor(Gender,
                           levels = c("0", "1"))) %>%
  mutate(clicks = factor(Clicks,
                           levels = c("1", "0"))) %>%
  glm(gender ~ gender * clicks,
      family = "binomial",
      data = .) %>%
  tidy()
new_click_data_results
```
