# Impute missingness

```{r load naniar, message= FALSE}

# load libraries
library(tidyverse)
library(naniar)
```

Create x, a vector, with values NA, NaN, Inf, ".", and "missing"

```{r make missing data, message= FALSE}

x <- c(NA, NaN, Inf, ".", "missing")

# Use any_na() and are_na() on to explore the missings
any_na(x)
are_na(x)

# Use n_miss() to count the total number of missing values in x
n_miss(x)

# Use n_complete() on dat_hw to count the total number of complete values
n_complete(x)
```

Use prop_miss() and prop_complete on dat_hw to count the total number of missing values in each of the variables:
```{r proportion of missingness, message= FALSE}

prop_miss(x)
prop_complete(x)
```


Use tarwars data to use to find missing values
summarizing missingness

```{r load starwars, message= FALSE}

starwars %>%
  miss_var_summary()

starwars %>%
  group_by(birth_year) %>%
  miss_var_summary()

# Return the summary of missingness in each case, by group
starwars %>%
  miss_case_summary() %>%
  head()

starwars %>%
  group_by(birth_year) %>%
  miss_case_summary() %>%
  head()

# Tabulate missingness in each variable and case of the dataset
starwars %>%
  miss_var_table() %>%
  head()

# Tabulate the missingness in each variable and in grouped_by
starwars %>%
  miss_case_table() %>%
  head()

starwars %>%
  group_by(birth_year) %>%
  miss_var_table() %>%
  head()

# Tabulate the missingness in each variable and in grouped_by
starwars %>%
  miss_case_table() %>%
  head()

starwars %>%
  group_by(birth_year) %>%
  miss_case_table() %>%
  head()
```

Other summaries of missingness
Calculate the summaries for each span of missingness

For each `birth_year` variable, calculate the run of missingness:

```{r calculate missing summaries, message= FALSE}

starwars %>%
  miss_var_run(var = birth_year)

# For each `birth_year` variable, calculate the run of missingness every 2 years
starwars %>%
  miss_var_span(var = birth_year, span_every = 2)

# Visualize missing values
vis_miss(starwars)

# Visualize and cluster all of the missingness 
vis_miss(starwars, cluster = TRUE)

# Visualize and sort the columns by missingness
vis_miss(starwars, sort_miss = TRUE)
```

Visualizing missing cases and variables
Visualize the number of missings in variables using `gg_miss_var()`:

```{r visualize missing data, message= FALSE}

gg_miss_var(starwars)
gg_miss_var(starwars, facet = birth_year)

# Explore the number of missings in cases using `gg_miss_case()` and facet by the variable `birth_year`
gg_miss_var(starwars)
gg_miss_case(starwars, facet = birth_year)

# Explore the missingness pattern using gg_miss_upset()
gg_miss_upset(starwars)

# Explore how the missingness changes across the birth_year variable using gg_miss_fct()
gg_miss_fct(starwars, fct = birth_year)

# and explore how missingness changes for a span of 2 years 
gg_miss_span(starwars, var =birth_year, span_every = 100, facet = mass) # also possible to facet
```



Explicitly missing values
Searching for and replacing missing values

```{r search and replace missing data, message= FALSE}

df <- tibble::tribble(
  ~name,           ~x,  ~y,              ~z,  
  "N/A",           "1",   "N/A",           "-100", 
  "N/A",           "3",   "Not Available", "-99",
  "N/A",         "NA",  "29",              "-98",
  "Not Available", "-99", "25",              "-101",
  "John Smith",    "-98", "28",              "-1")


# Explore the strange missing values "N/A"
miss_scan_count(data = df, search = list("N/A"))

# Explore the strange missing values "missing"
miss_scan_count(data = df, search = list("missing"))

# Explore the strange missing values "na"
miss_scan_count(data = df, search = list("na"))

# Explore the strange missing values "Not Available"
miss_scan_count(data = df, search = list("Not Available"))

# Explore the strange missing values " " (a single space)
miss_scan_count(data = df, search = list(" "))

# Explore all of the strange missing values, "N/A", "missing", "na", " "
miss_scan_count(data = df, search = list("N/A", "missing","na", " ", "Not Available"))

# Replace the strange missing values "N/A", "na", and "missing" with `NA` for the variables x and y
df_clean <- replace_with_na(df, replace = list(x = c("N/A", "missing","na", " ", "Not Available"),
                                                       y = c("N/A", "missing","na", " ", "Not Available")))
```

Test if `df_clean` still has these values in it?

```{r df clean, message= FALSE}

miss_scan_count(df_clean, search = list("N/A", "missing","na", " ", "Not Available"))

# Use `replace_with_na_at()` to replace with NA
replace_with_na_at(df,
                   .vars = c("x", "y", "name"), 
                   ~.x %in% c("N/A", "missing","na", " ", "Not Available"))

# Use `replace_with_na_if()` to replace with NA the character values using `is.character`
replace_with_na_if(df,
                   .predicate = is.character, 
                   ~.x %in% c("N/A", "missing","na", " ", "Not Available"))

# Use `replace_with_na_all()` to replace with NA
replace_with_na_all(df, ~.x %in% c("N/A", "missing","na", " ", "Not Available"))

```


Implicitly missing values

Print the frogger data to have a look at it
starwars

Use `complete()` on the `time` variable to make implicit missing values explicit

```{r implicit NA, message= FALSE}

starwars_tidy <- starwars %>% complete(name, hair_color)

# Use `fill()` to fill down the name variable in the starwars dataset
starwars %>% fill(name)

# Correctly fill() and complete() missing values so that our dataset becomes sensible
starwars %>%
  fill(name) %>%
  complete(name, hair_color)

# Missing Data dependence
# Exploring missingness dependence
# Arrange by birth_year
starwars %>% arrange(birth_year) %>% vis_miss() # this variable is important for missingness

# Arrange by mass
starwars %>% arrange(mass) %>% vis_miss() # this variable is important for missingness

# Arrange by height 
starwars %>% arrange(height) %>% vis_miss() # this variable is important for missingness
```


Tools to explore missing data dependence
data to use
airquality

Create shadow matrix data with `as_shadow()`

```{r tools for missing data, message= FALSE}

as_shadow(airquality) %>%
  head()

# Create nabular data by binding the shadow to the data with `bind_shadow()`
bind_shadow(airquality) %>%
  head()

# Bind only the variables with missing values by using bind_shadow(only_miss = TRUE)
bind_shadow(airquality, only_miss = TRUE) %>%
  head()
```

performing group summaries
`bind_shadow()` and `group_by()` humidity missingness (`Ozone_NA`)

```{r group summaries, message= FALSE}

airquality %>%
  bind_shadow() %>%
  group_by(Ozone_NA) %>%
  summarise(Wind_mean = mean(Wind), # calculate mean of Wind
            Wind_sd = sd(Wind)) # calculate standard deviation of Wind

# Repeat this, but calculating summaries for Temp
airquality %>%
  bind_shadow() %>%
  group_by(Ozone_NA) %>%
  summarise(Temp_mean = mean(Temp),
            Temp_sd = sd(Temp))


# Summarise wind by the missingness of `Solar.R_NA`
airquality %>% 
  bind_shadow() %>%
  group_by(Solar.R_NA) %>%
  summarise(Wind_mean = mean(Wind), # calculate mean of Wind
            Wind_sd = sd(Wind), # calculate standard deviation of Wind
            n_obs = n())

# Summarise wind by missingness of `Solar.R_NA` and `Ozone_NA`
airquality %>%
  bind_shadow() %>%
  group_by(Solar.R_NA, Ozone_NA) %>%
  summarise(Temp_mean = mean(Temp),
            Temp_sd = sd(Temp),
            n_obs = n())

```

Visualizing missingness across one variable
First explore the missingness structure of `airquality` using `vis_miss()`

```{r visualize across variables, message= FALSE}

vis_miss(airquality)

# Explore the distribution of `Temp` for the missingness of `Ozone_NA` using  `geom_density()`
bind_shadow(airquality) %>%
  ggplot(aes(x = Temp, 
             color = Ozone_NA)) + 
  geom_density()

# Explore the distribution of Temp for the missingness of humidity (Solar.R_NA) using  `geom_density()`
bind_shadow(airquality) %>%
  ggplot(aes(x = Temp,
             color = Solar.R_NA)) + 
  geom_density()

# Explore the distribution of Wind for the missingness of Solar.R_NA using  `geom_density()` and facetting by the missingness of (`Solar.R_NA`).
airquality %>%
  bind_shadow() %>%
  ggplot(aes(x = Wind)) + 
  geom_density() + 
  facet_wrap(~Solar.R_NA)

# Build upon this visualisation by filling by the missingness of (`Solar.R_NA`).
airquality %>%
  bind_shadow() %>%
  ggplot(aes(x = Wind,
             color = Solar.R_NA)) + 
  geom_density() + 
  facet_wrap(~Solar.R_NA)


# Explore the distribution of wind for the missingness of air temperature using  `geom_boxplot()`
airquality %>%
  bind_shadow() %>%
  ggplot(aes(x = Ozone_NA,
             y = Wind)) + 
  geom_boxplot()

# Build upon this visualisation by facetting by the missingness of Ozone_NA 
airquality %>%
  bind_shadow() %>%
  ggplot(aes(x = Ozone_NA,
             y = Wind)) + 
  geom_boxplot() + 
  facet_wrap(~Ozone_NA)
```


Visualizing missingness across two variables

Explore the missingness in wind and air temperature, and display the missingness using `geom_miss_point()`:

```{r visualize 2 missing vars, message= FALSE}

ggplot(airquality,
       aes(x = Wind,
           y = Temp)) + 
  geom_miss_point()

# Explore the missingness in Ozone and Solar.R, and display the missingness using `geom_miss_point()`
ggplot(airquality,
       aes(x = Ozone,
           y = Solar.R)) + 
  geom_miss_point()


# Explore the missingness in Ozone and Solar.R, and display the missingness using `geom_miss_point()`. Facet by Temp to explore this further.
ggplot(airquality,
       aes(x = Ozone,
           y = Solar.R)) + 
  geom_miss_point() + 
  facet_wrap(~Temp)

# Use geom_miss_point() and facet_wrap to explore how the missingness in Ozone  is different for missingness of Solar.R
bind_shadow(airquality) %>%
  ggplot(aes(x = Ozone,
             y = Solar.R)) + 
  geom_miss_point() + 
  facet_wrap(~Temp)

# Use airquality() and facet_grid to explore how the missingness in wind_ew and air_temp_c is different for missingness of Temp AND by Month- by using `facet_grid(Temp ~ Month)`
bind_shadow(airquality) %>%
  ggplot(aes(x = Ozone,
             y = Solar.R)) + 
  geom_miss_point() + 
  facet_grid(Temp ~ Month)
```


Filling in the blanks
Impute the data below the range using `impute_below`.

```{r fill in blanks, message= FALSE}

airquality_imp <- impute_below_all(airquality)

# Visualise the new missing values
ggplot(airquality_imp,
       aes(x = Wind, y = Temp)) + 
  geom_point()

# Impute and track data with `bind_shadow`, `impute_below`, and `add_label_shadow`
airquality_imp_track <- bind_shadow(airquality) %>% 
  impute_below_all() %>%
  add_label_shadow()

# Look at the imputed values
airquality_imp_track
```

Visualise imputed values in a scatterplot
Visualise the missingness in wind and air temperature, coloring missing values with Ozone_NA:

```{r scatterplot of missings, message= FALSE}

ggplot(airquality_imp_track,
       aes(x = Wind, y = Temp, color = Ozone_NA)) + 
  geom_point()

# Visualise humidity and air temp, coloring any missing cases using the variable any_missing
ggplot(airquality_imp_track,
       aes(x = Wind, y = Temp, color = any_missing)) + 
  geom_point()
```

Create histogram of imputed data
Explore the values of air_temp_c, visualising the amount of missings with `air_temp_c_NA`.

```{r histogram of missings, message= FALSE}

p <- ggplot(airquality_imp_track, aes(x = Ozone, fill = Ozone_NA)) + geom_histogram()

# Expore the missings in humidity using humidity_NA
p2 <- ggplot(airquality_imp_track, aes(x = Solar.R, fill = Solar.R_NA)) + geom_histogram()

# Explore the missings in airquality_imp_track according to Month, using `facet_wrap(~Month)`.
p + facet_wrap(~Month)

# Explore the missings in humidity according to Month, using `facet_wrap(~Month)`.
p2 + facet_wrap(~Month)
```


What makes a good imputation
Impute the mean value and track the imputations 

```{r impute, message= FALSE}

airquality_imp_mean <- bind_shadow(airquality) %>% 
  impute_mean_all() %>% 
  add_label_shadow()

# Explore the mean values in wind in the imputed dataset
ggplot(airquality_imp_mean, 
       aes(x = Ozone_NA, y = Wind)) + 
  geom_boxplot()

# Explore the values in air wind in the imputed dataset
ggplot(airquality_imp_mean, 
       aes(x = Solar.R_NA, y = Wind)) + 
  geom_boxplot()

# Explore imputations in air temperature and wind, coloring by the variable, any_missing
ggplot(airquality_imp_mean, 
       aes(x = Wind, y = Temp, color = any_missing)) + 
  geom_point()

# Explore imputations in air temperature and wind, coloring by the variable, any_missing, and faceting by Month
ggplot(airquality_imp_mean, 
       aes(x = Wind, y = Temp, color = any_missing)) +
  geom_point() + 
  facet_wrap(~ Month)


# Gather the imputed data 
airquality_imp_mean_gather <- shadow_long(airquality_imp_mean,
                                     Temp,
                                     Wind)
# Inspect the data
airquality_imp_mean_gather

# Explore the imputations in a histogram 
ggplot(airquality_imp_mean_gather, 
       aes(x = as.numeric(value), fill = value_NA)) + 
  geom_histogram() + 
  facet_wrap(~ variable) 
```


Performing imputations
Impute Temp and Wind, and track missing values

```{r simputation library, message= FALSE}

library(simputation)

airquality_imp_lm_wind <- airquality %>% 
  bind_shadow() %>%
  impute_lm(Temp ~ Wind + Month) %>% 
  impute_lm(Temp ~ Wind + Day) %>%
  add_label_shadow()

# Plot the imputed values for Temp and Wind, colored by missingness
ggplot(airquality_imp_lm_wind, 
       aes(x = Temp, y = Wind, color = any_missing)) +
  geom_point()

# Evaluating and comparing imputations
## Bind the models together 
bound_models <- bind_rows(mean = airquality_imp_mean,
                          lm_wind = airquality_imp_lm_wind,
                          .id = "imp_model")

# Inspect the values of air_temp and humidity as a scatterplot
ggplot(bound_models,
       aes(x = Wind,
           y = Temp,
           color = any_missing)) +
  geom_point() + 
  facet_wrap(~imp_model)


# Build a model adding Month to the outcome
airquality_imp_lm_wind_month <- bind_shadow(airquality) %>%
  impute_lm(Temp ~ Wind + Day + Month) %>%
  impute_lm(Temp ~ Wind + Ozone + Month) %>%
  add_label_shadow()

# Bind the mean, lm_wind, and lm_wind_year models together
bound_models <- bind_rows(mean = airquality_imp_mean,
                          lm_wind = airquality_imp_lm_wind,
                          lm_wind_year = airquality_imp_lm_wind_month,
                          .id = "imp_model")

# Explore air_temp and humidity, coloring by any missings, and faceting by imputation model
ggplot(bound_models, aes(x = Wind, y = Temp, color = any_missing)) +
  geom_point() + facet_wrap(~imp_model)


# Gather the data and inspect the distributions of the values
bound_models_gather <- bound_models %>%
  select(Temp, Wind, any_missing, imp_model) %>%
  gather(key = "key", value = "value", -any_missing, -imp_model)

# Inspect the distribution for each variable, for each model
ggplot(bound_models_gather,
       aes(x = imp_model, y = value, color = imp_model)) +
  geom_boxplot() + facet_wrap(~key, scales = "free_y")

# Inspect the imputed values
bound_models_gather %>%
  filter(any_missing == "Missing") %>%
  ggplot(aes(x = imp_model, y = value, color = imp_model)) +
  geom_boxplot() + facet_wrap(~key, scales = "free_y")
```


Combining and comparing many imputation models
Create an imputed dataset using a linear models

```{r compare imputation models, message= FALSE}

airquality_imp_lm_all <- bind_shadow(airquality) %>%
  add_label_shadow() %>%
  impute_lm(Temp ~ Wind + Day + Month) %>%
  impute_lm(Temp ~ Wind + Ozone + Month) 

# Bind the datasets
cc = airquality
imp_lm_wind = as.data.frame(airquality_imp_lm_wind) %>%
  mutate_if(is.factor, ~ as.numeric(levels(.x))[.x])

imp_lm_all = as.data.frame(airquality_imp_lm_all) %>%
  mutate_if(is.factor, ~ as.numeric(levels(.x))[.x])

bound_models <- bind_rows(airquality,
                          imp_lm_wind,
                          imp_lm_all,
                          .id = "imp_model")

# Look at the models
head(bound_models,4)

# Create the model summary for each dataset
library(broom)
library(tidyr)
model_summary <- bound_models %>% 
  group_by(imp_model) %>%
  nest() %>%
  mutate(mod = map(data, ~lm(Temp ~ Wind + Ozone + Month, data = .)),
         res = map(mod, residuals),
         pred = map(mod, predict),
         tidy = map(mod, tidy))

# Explore the coefficients in the model
model_summary %>% 
  select(imp_model,tidy) %>%
  unnest()

best_model <- "imp_lm_all"
```
