---
name: Noushin Nabavi, PhD
position: "Data Science Fellow"
address: "Victoria, BC"
phone: +1 778-999-1982
www: noushinn.GITHUB.IO
email: "nabavi.noushin@gmail.com"
twitter: nabavinoushin
github: noushinn
linkedin: noushin-nabavi-56b27b13
date: "`r format(Sys.time(), '%B %Y')`"
output: vitae::awesomecv
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
library(tidyverse)
```

# Career Highlights

•	Results-driven, forward-thinking accomplished Research Scientist
•	Demonstrated leadership and project management experience in private and public sectors
•	Skilled quantitative analyst proficient with Python, R and Microsoft Office 
•	Experienced delivery of products including wrangled data, analytics concepts and documentation: from rigorously defensible definitions through to implementation and evaluation 
•	Service-oriented team player dedicated to delivering quality services and gaining the trust of clients and stakeholders


# Education
```{r}
tribble(
  ~ degree, ~ uni, ~ loc, ~ dates, ~ details,
  "Certifications", "Data Camp", "Online", "2018-2019", "Data Analytics using R",
  "Certifications", "Mitacs Inc.", "University of British Columbia", "2017-2019", "Project Management and Agile Transformation",
  "Certifications", "Yale Centre for Integration of Teaching and Learning", "Online Summer Institute for Teaching", "2018", "Teaching best practices",
  "mini-MBA", "University of British Columbia", "Vancouver, BC, Canada", "2018", "Business Administration for Academics",
  "Doctor of Philosophy", "University of Toronto", "Toronto, ON, Canada", "2005-2011", "Cell and Systems Biology",
  "Bachelor of Science", "University of Toronto", "Toronto, ON, Canada", "2001-2005", "Honors in Cell and Molecular Biology"
  
) %>% 
  detailed_entries(degree, dates, uni, loc, details)
```

# Professional Experience
```{r}
tribble(
  ~ role, ~ company, ~ loc, ~ dates, ~ details,
  "Mitacs Science Policy Fellow", "BC Government", "Victoria, BC, Canada", "September 2018 - September 2019", "Analyze and communicate complex administrative integrated data for evidence based policy making; Build and strengthen effective stakeholder relationships with ministry clients and contractors; Manage a data science project while developing and implementing timely methodologies and performance metrics",
  
  "Senior Research Fellow", "University of British Columbia", "Vancouver, BC, Canada", "February 2015 - February 2018", "Assemble and lead an international consortium of scientists for a genomics project, published 5 manuscripts and presented the results in 3 international conferences and 2 local symposia; Analyze and integrate genomics, transcriptomics, and proteomics datasets using R and proprietary software; Wrote and secured a $275K operating grant from BC Cancer Foundation for next generation sequencing tumor profiling of malignant peritoneal mesothelioma",

  "Postdoctoral Fellow", "University of California, San Francisco", "San Francisco, CA, USA", "August 2013 - December 2014", "Design research project and conduct experiments related to liver disease with an expert team of scientists; Analyze experimental data and integrate results to generate presentations and publications; Contribute to securing a $427K NIH grant",

    "Postdoctoral Fellow", "University of California, Berkeley", "Berkeley, CA, USA", "January 2012 - July 2013", "Implement a reproducible scientific workflow for a project related to aging: designed and carried out experiments with animal models for aging, analyzed results; Manage a research project with graduate students and undergraduate students towards publication in a peer-review journal",

  "Teaching Assistant and Research Co-Supervisor", "University of Toronto", "Toronto, ON, Canada","September 2005 - December 2011", "Rank the top 10% among departmental teaching assistants for 500+ hours of instruction in introductory biology laboratory, cellular molecular biology, and microbiology; Create and delivere departmental lectures to 500+ students with satisfactory responses from students; Co-supervise 10 undergraduate thesis students on successful thesis projects maintaining 100% retention rate (8 students secured successful science jobs, 2 pursued graduate studie)"
  
) %>% 
  detailed_entries(role, dates, company, loc, details)
```

# Awards & Achievements
## Awards
```{r}
tribble(
  ~ award, ~ from, ~ year,
"Premier’s Award in Evidence-Based Design", "BC Gov, Finalist", "2019",  
"Travel award", "Hackseq Genomics Hackathon", "2018-2019",
"Mentorship award", "Victoria Health Hackathon", "2018",
"Mitacs Canadian Science Policy Fellowship", "Mitacs Inc.", "2018",
"BeHeard CRISPR Technology", "Rare Genomics Institute", "2017",
"Amazon Web Services for cloud computing", "Amazon", "2016",
"Mitacs Accelerate Fellowship", "Mitacs Inc.", "2015",
"Science Teaching Award", "American Society for Microbiology", "2014",
"Siebel Stem Cell Fellowship", "Siebel Foundation", "2012",
"Graduate Student Research Award", "University of Toronto", "2010",
"Graduate Student Research", "Natural Science and Engineering Research Council", "2009",
"Women in Science and Mathematics", "General Motors of Canada", "2008",
"Travel awards", "American Society for Cell Biology", "2006-2007"

) %>%
  brief_entries(award, year, from)
```

# Selected talks
```{r}
tribble(
  ~name, ~event, ~location, ~year,
  
"Healing old wounds", "Institute of Public Administration of Canada", "Victoria, BC, Canada", "October 2018",
"Reproducible data cleaning workflow with R", "Data Science Community of Practice", "Victoria, BC, Canada", "May 2019", 
"BAP1 creates a potential therapeutic vulnerability in malignant peritoneal mesothelioma", "International Mesothelioma Interest Group", "Ottawa, ON, Canada", "May 2018",
"Comprehensive molecular characterization of peritoneal mesothelioma", "Vancouver Prostate Centre Friday Seminar", "Vancouver, BC, Canada", "February 2017",
"Cytochrome P450 degradation in hepatocytes and implications in drug metabolism","UCSF Postdoctoral Symposium", "San Francisco, CA, USA",  "June 2014",
"SIRT2’s role in stem cell biology", "Berkeley Stem Cell Center Retreat, Fred Farr Forum", "Asilomar, CA, USA", "April 2012 and 2013"
  
) %>% 
  detailed_entries(
    what = name,
    when = year,
    with = event,
    where = location
  )
```